---
title: "General information about GAMs"
author: "Pablo E. Gutiérrez-Fonseca"
date: "3/18/2022"
output:
  pdf_document: default
---

## **Model construction** 

### **s (Both GAMs)**
s = represent smooth function
```{r, eval=FALSE}
gam(value ~s(date,...
```


### **Define knots (Both GAMs)**
k = knots. 12 month per year or 24 sampling event per year.\
Seleccione 12 por Simpson, del siguiente enlace:\
https://fromthebottomoftheheap.net/2014/05/09/modelling-seasonal-data-with-gam/


### **bs= basis spline (Both GAMs)**
bs= basis spline.\
Smooth classes are invoked directly by s terms.\
https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/smooth.terms.html

```{r, eval=FALSE}
brms::brm(bf(value ~ s(date, bs="cs"...))
```

#### **Cubic regression splines & A cyclic cubic regression spline.**\
bs="cr". These have a cubic spline basis defined by a modest sized set of knots spread evenly through the covariate values.\
bs="cs" specifies a shrinkage version of "cr".\
bs="cc" specifies a cyclic cubic regression splines. i.e. a penalized cubic regression splines whose ends match, up to second derivative.\


#### **P-splines**\
bs="ps".

#### **bf (Bayesian GAMs)**
Note that we use the bf() argument to specify this nonlinear model.


## **Output**

Hay que ver el Smooth Terms: ->   sds(sdate_1) -> 
sds(stimes_1) is the variance parameter, which has the effect of controlling 
the wiggliness of the smooth - the larger this value the more wiggly the smooth.
https://fromthebottomoftheheap.net/2018/04/21/fitting-gams-with-brms/


## **Check models**

### pp_check()
The pp_check allows for graphical posterior predictive checking. We can generate figures to compare the observed data to simulated data from the posterior predictive distribution. This is a great 
graphical way to evaluate your model.\
https://tem11010.github.io/regression_brms/

Here, nsamples refers to the number of draws from the posterior distribution to use to calculate yrep values.

```{r, eval=FALSE}
pp_check(model, nsamples=100)
```


### **bayes_r2**

Bayes R2 quantifies the expected fit or variance explained by a model

We can also get an R-squared estimate for our model, 
thanks to a newly-developed method from Andrew Gelman, 
Ben Goodrich, Jonah Gabry and Imad Ali, with an explanation here.
http://www.stat.columbia.edu/~gelman/research/unpublished/bayes_R2.pdf
https://tem11010.github.io/regression_brms/

r2(cc.qp_A.Bayes_mod) Existe esta otra, pero usare la de Gelman

```{r, eval=FALSE}
bayes_R2(model)
```

### **loo_compare()**

The loo_compare() output rank orders the models such that the best fitting model appears on top. All models receive a difference score relative to the best model.
https://bookdown.org/ajkurz/DBDA_recoded/model-comparison-and-hierarchical-modeling.html


### **brms::model_weights()**

I don't know that I'd call these weights probabilities, but they do sum to one.
https://bookdown.org/ajkurz/DBDA_recoded/model-comparison-and-hierarchical-modeling.html



### **AIC**

AIC and DIC are not recommended in particular not with Bayesian models. So it is not necessarily surprising that the results differ (buerkner).\

https://discourse.mc-stan.org/t/waic-aicc-and-dic-differences-in-a-bernoulli-glmm-in-brms-vs-glmmtmb/5966

Gelman, A., Hwang, J., & Vehtari, A. (2014). Understanding predictive information criteria for Bayesian models. Statistics and computing, 24 (6), 997-1016.

Vehtari, A., Gelman, A., & Gabry, J. (2017). Practical Bayesian model evaluation using leave-one-out cross-validation and WAIC. Statistics and Computing , 27 (5), 1413-1432.
